---
title: "Data 607_Hw9_Web APIs"
author: "Sin Ying Wong"
date: "10/27/2019"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
    extra_dependencies: ["geometry", "multicol", "multirow"]
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Instruction

The New York Times web site provides a rich set of APIs, as described here: <https://developer.nytimes.com/apis>(https://developer.nytimes.com/apis) 

You’ll need to start by signing up for an API key. 

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

# Load packages

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(stringr)
```


```{r api, echo=FALSE}
api_key <- 'I005E9HTTpYWKlx2Vztc1DiCS01pll1G'

```

# Read data

Searching for Climate articles released in the first five months in 2019

```{r read_data}
url = str_c("https://api.nytimes.com/svc/search/v2/articlesearch.json", 
            '?fq=news_desk:("Climate")',
            '&begin_date=',"20190101",
            '&end_date=',"20190531",
            '&api-key=',api_key,
            sep = '')

url

data <- GET(url)

data

http_status(data)

stop_for_status(data)
```

# Convert to DataFrame

```{r dataframe}
climate_df <- fromJSON(url) %>% 
  .$response %>% 
  .$doc %>%
  mutate(headline = headline$main,
         pub_date = as.Date(pub_date),
         byline = byline$original) %>%
  mutate(byline = str_remove(byline, '^By ')) %>%
  rename(author = byline, 
         date = pub_date) %>%
  select(date,
         type_of_material,
         headline,
         abstract,
         author,
         web_url)

climate_df

```

